---
title: "Diets Part 1: Ontogenetic diet preferences from REEM"
author: "Alberto Rovellini"
date: "1/12/2022"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# Purpose

Here, we calculate weight-based diet preferences of GOA predators based on bottom trawl data. This document characterizes ontogenetic preferences based on size at 50% maturity. Data is diet data from the REEM group [available here](https://apps-afsc.fisheries.noaa.gov/refm/reem/webdietdata/dietdataintro.php).

The goal is to define generic diet prefernces for each species, to be used as PPREY entries in Atlantis GOA. We use diet data from 1990 to 2013 to avoid capturing the heath wave, when diets shifted substantially. Remember, these will only be the initial, pre-calibration diet preferences in Atlantis.

The workflow is as follows:

1. Read in the diet and prey length data from  REEM.
2. Read in information on size at 50% maturity for each (fish) species, pulling from FishLife.
3. Read in information on length-weight relationships, pulling from Froese et al. (2014).
4. Consider the predator data set. Assign all predator individuals (that we can based on available lmat) to juveniles and adults based on length at 50% maturity. Then at haul level, for predator $i$ and prey $j$, calculate the weight of $j$ eaten by $i$. Set this aside for later. We will use information from the stage composition calculated below to break down the weight of $j$ between juveniles and adults. 
5. Consider the prey-length data set. For a subset of the fish sampled for diet analysis (which were in turn already a subset of the fish caught in a haul), the length of a subset of the prey items in their stomach is measured. This contains a column named Prey_sz1, which is "An 8.3-digit numeric field that does not allow null values representing the standard length (mm) of fish prey, carapace width (mm) of Majid crabs, or carapace length of Lithodid crabs. This is the primary prey size measurement". Attribute each predator to a life stage (juvenile or adult) based on the size at 50% maturity. Do the same for the prey, where possible (the invertebrates are not split into life stages for our purposes).
6. Within each predator, turn the length of each prey item to a weight based on the length-weight relationship $W=a*L^b$.
7. Group by **stratum**, predator, and life stage, and sum up the weight of each prey species by life stage, and work out the proportions of the prey of each life stage ingested (e.g. adults of this predator in this haul ate 1000 kg of this prey, 20% in juveniles and 80 as adults, by weight).
8. At this point you should have, for each stratum, each species, each life stage, and each prey, a proportion of juvenile and one for adult prey. Use these proportions to break down the weight of prey to juveniles and adults **at haul level**. 
9. Apply the steps from diet_exploration_atlantis_groups_all_indeces.Rmd to calculate weight-based preference for a prey species, weighted by the number of predators **with a full stomach in a haul**. Do this 4 times for all the combination of juvenile and adult predators and prey. This should yield four matrices for vertebrates. Biomass pools will only have two matrices. At this stage, before calculating the proportions, map the weight of ingested prey to Atlantis groups by adding appropriately. 

Bear in mind that we will not be able to examine all predator-prey combinations from the data. Treat absences as zeroes in this case (e.g. no juvenile cods were found in ATF gut contents? Then ATF only eats adult cods).

```{r}
library(tidyverse)
library(viridis)
library(FishLife)
library(rfishbase)
```

```{r}
select <- dplyr::select
```

# Read data (1-3)

## REEM data

```{r}
predator <- read.csv("../data/1_Gulf Of Alaska_RawPP.csv")
prey_length <- read.csv("../data/1_Gulf Of Alaska_RawPL.csv")

# clean column names
prey_length <- prey_length %>% set_names(gsub("X.","",colnames(prey_length)))

# drop "Dasycottus setiger(spinyhead sculpin)" since there is one, it is not useful, and it causes issues later due to a typo
prey_length <- prey_length %>% filter(Pred_name!='Dasycottus setiger(spinyhead sculpin)')

# limit to 1990 onward - older data is tricky
predator <- predator %>% filter(between(Year, 1990, 2011)) # leaving out the heat wave years
prey_length <- prey_length %>% filter(between(Year, 1990, 2011)) # leaving out the heat wave years

# read in haul information
hauls <- read.csv("../data/Haul Descriptions.csv", fileEncoding = "UTF-8-BOM")

# add stratum definition to hauls, in both predator and prey set
# note: Hauljoin is (surprisingly) different between RACE-GAP and REEM data. Joining by vessel, cruise, and haul seem to work though
predator <- predator %>% 
  left_join(hauls %>% select(Vessel.Number,Cruise.Number,Haul.Number,Stratum), by = c('Vessel'='Vessel.Number','Cruise'='Cruise.Number','Haul'='Haul.Number')) %>% 
  filter(!is.na(Stratum))

prey_length <- prey_length %>% 
  left_join(predator %>% select(Hauljoin,Stratum) %>% distinct(), by='Hauljoin') %>% 
  filter(!is.na(Stratum))
```

What species can we carry out the size structure analysis for? The limiting set will be the prey length set. We will need a species list for predator and prey, but take both from the prey set as that is the smallest.
```{r}
pred_species <- prey_length %>% 
  rowwise() %>%
  mutate(Scientific_name = paste(word(Pred_name, 1:2), collapse = ' ')) %>% 
  ungroup() %>%
  select(Scientific_name) %>% 
  distinct() %>% 
  arrange() %>%
  mutate(tt = strsplit(Scientific_name, ' ')) %>%
  rowwise() %>%
  mutate(Genus = tt[1],
         Species = tt[2]) %>%
  select(Genus, Species) %>%
  drop_na() %>%
  ungroup()

# prey list is a lot messier 
prey_species <- prey_length %>% 
  rowwise() %>%
  mutate(Prey_Name = str_replace(Prey_Name,'\\(',''),
         Prey_Name = str_replace(Prey_Name,'\\)',''),
         Scientific_name = ifelse(str_count(Prey_Name, '\\w+')>1, 
                                  paste(word(Prey_Name, (str_count(Prey_Name, '\\w+')-1):str_count(Prey_Name, '\\w+')), collapse = ' '),
                                  Prey_Name)) %>% 
  ungroup() %>%
  select(Scientific_name,Prey_nodc) %>% 
  distinct() %>% 
  arrange() %>%
  mutate(tt = strsplit(Scientific_name, ' ')) %>%
  rowwise() %>%
  mutate(Genus = tt[1],
         Species = tt[2]) %>%
  select(Genus, Species) %>%
  drop_na() %>%
  ungroup()
# all those ending with sp. or with common names will not work, as well as a few other with dashes etc. However, those will likely not have info anyway. There are a lot of invertebrates here too, which will also not work.
```

## Length at 50% maturity

For the purpose of splitting between juveniles and adults, we use estimates of length at maturity from the `FishLife` [package](https://github.com/James-Thorson-NOAA/FishLife). 

This needs to be done for the predator and prey list both. 
```{r}
source('funs_lw_lmat.R')

# add pollock, it still goes by its old name in Fishbase for some reason
pred_species <- rbind(pred_species, c('Theragra','chalcogramma'))
prey_species <- rbind(prey_species, c('Theragra','chalcogramma'))

# these are needed in the call to get_lmat
fl_params <- (FishLife::FishBase_and_RAM$beta_gv) # all species params
fl_names <- row.names(fl_params)
fl_covmat <- (FishLife::FishBase_and_RAM$Cov_gvv) # all species variance covariance

lmat_pred <- pred_species %>% 
  mutate(lmat = purrr::pmap(list(this_genus=Genus,this_species=Species),get_lmat)) %>%
  unnest(cols = c(lmat), keep_empty = T)

lmat_prey <- prey_species %>% 
  mutate(lmat = purrr::pmap(list(this_genus=Genus,this_species=Species),get_lmat)) %>%
  unnest(cols = c(lmat), keep_empty = T)
```

## Length-weight relationship

Taking parameters from [Froese et al. (2014)](https://doi.org/10.1111/jai.12299). The form of the LW relationship in the Bayesian hierarchical analysis is $W=a*L^b$. I apply their code in `SingleSpeciesLWR_7.R` to obtain mean estimates of log10a and b from Bayesian hierarchical analysis based on (1) priors that are specific to body shape of the fish, and (2) previous studies that act as data to get a posterior distribution of the parameters. Importantly, for these parameters, weight is in grams and length is in centimeters.
```{r}
lw_pred <- pred_species %>% 
  mutate(lw_bayes = purrr::pmap(list(this_genus=Genus,this_species=Species),possibly(apply_Froese,NA))) %>%
  unnest_wider(col = c(lw_bayes)) %>%
  set_names(c('Genus','Species','Log10a','b')) %>%
  mutate(a=10^Log10a) %>% # because the models return log10a, not a
  select(-Log10a)

lw_prey <- prey_species %>% 
  mutate(lw_bayes = purrr::pmap(list(this_genus=Genus,this_species=Species),possibly(apply_Froese,NA))) %>%
  unnest_wider(col = c(lw_bayes)) %>%
  set_names(c('Genus','Species','Log10a','b')) %>%
  mutate(a=10^Log10a) %>% # because the models return log10a, not a
  select(-Log10a)
```

Join length-weight and length at maturity keys. 
```{r}
lw_lmat_pred <- lw_pred %>% 
  left_join(lmat_pred, by = c('Genus','Species')) %>%
  drop_na() %>%
  mutate(Scientific.Name = paste(Genus, Species, sep = ' ')) %>%
  select(-Genus,-Species)

lw_lmat_prey <- lw_prey %>% 
  left_join(lmat_prey, by = c('Genus','Species')) %>%
  drop_na() %>%
  mutate(Scientific.Name = paste(Genus, Species, sep = ' ')) %>%
  select(-Genus,-Species)

# restore pollock to the correct name
lw_lmat_pred$Scientific.Name <- gsub('Theragra chalcogramma','Gadus chalcogrammus',lw_lmat_pred$Scientific.Name)
lw_lmat_prey$Scientific.Name <- gsub('Theragra chalcogramma','Gadus chalcogrammus',lw_lmat_prey$Scientific.Name)
```

# The predator set (4)

Point 4. Take the predator set and the predator length at maturity information, and assign each predator individual to juvenile and adult based on lmat. Then, at haul level, for each life stage of each predator add up the weight of each prey species (pooled regardless of life stage for now).
```{r}
assign_stage <- function(Name,Len,lw_lmat){
  this_lmat <- lw_lmat %>% filter(Scientific.Name == Name) %>% pull(lmat)
  this_stage <- ifelse(length(this_lmat)==0, 'U', ifelse(Len<this_lmat,'J','A')) # if not possible to assign, use U for undetermined
  this_stage
}

predator <- predator %>% 
  mutate(Pred_stage = purrr::pmap(list(Pred_name,Pred_len),possibly(assign_stage,NA),lw_lmat_pred)) %>%
  unnest(cols = Pred_stage)
```
The predators that do not have a value for A or J are not present as predators in the "prey_length" set. Meaning that none of them has ever been examined for ingested prey length, and we will not be able to identify stage-specific trophic relationships for them.

Set up a counter for the n of predators with a non-empty stomach in each haul. This becomes a property of the haul like the hauljoin field and it will be used later to weight diet comp proportions between hauls based on the number of stomachs in a haul, in such a way that hauls containing more samples are given more importance. Then group the predator set by year, haul (with stratum), predator name, **predator stage**, and prey name, and sum up the prey weight ingested in that haul by that predator. 
```{r}
stomach_counter <- predator %>%
  filter(Pred_stomwt>0) %>% # use only non-empty stomachs
  group_by(Year,Hauljoin,Stratum,Pred_name,Pred_stage) %>% 
  mutate(Npred = length(unique(Pred_specn))) %>% 
  ungroup() %>%
  select(Hauljoin,Pred_name,Pred_stage,Npred) %>%
  distinct()

predator_by_haul <- predator %>%
  left_join(stomach_counter, by = c('Hauljoin','Pred_name','Pred_stage')) %>%
  group_by(Year,Hauljoin,Stratum,Npred,Pred_name,Pred_nodc,Pred_stage,Prey_Name,Prey_nodc) %>%
  summarise(Tot_weight_ingested = sum(Prey_twt)) %>% 
  ungroup()
```
Set this aside for later.

# The prey set (5-7)

This is going to be painful to deal with if nothing else for the names, which for some reason are in a different format from the other data set, making a join impossible. Alter the names based on how we extracted species names above.
```{r}
prey_length1 <- prey_length %>% 
  rowwise() %>%
  mutate(Prey_sz_cm = Prey_sz1/10, # size from mm to cm to apply length-weight transformations appropriately
         Pred_name_new = paste(word(Pred_name, 1:2), collapse = ' '),
         Prey_Name = str_replace(Prey_Name,'\\(',''),
         Prey_Name = str_replace(Prey_Name,'\\)',''),
         Prey_name_new =  ifelse(str_count(Prey_Name, '\\w+')>1, 
                                 paste(word(Prey_Name, (str_count(Prey_Name, '\\w+')-1):str_count(Prey_Name, '\\w+')), collapse = ' '),
                                 Prey_Name)) %>%
  ungroup() # NOTE: this returns imperfect names, in instances where the scientific name was not provided / incomplete (see above). Keep an eye on how this propagates

# assign each predator and prey to their stage
prey_length1 <- prey_length1 %>%
  mutate(Pred_stage = purrr::pmap(list(Pred_name_new,Pred_len),possibly(assign_stage,NA),lw_lmat_pred),
         Prey_stage = purrr::pmap(list(Prey_name_new,Prey_sz_cm),possibly(assign_stage,NA),lw_lmat_prey)) %>%
  unnest(cols = c(Pred_stage, Prey_stage)) 
```

Turn the length of each prey to a weight with the length-weight key (where possible) and the relationship $W=a*L^b$. Have NA where not possible.

**NOTE**: 

1. I checked the LW curves defined by the a and b parameters that we pull this way and they seem to make sense in terms of unites being cm and g for length and weight, respectively. Important to note that here we use estimates from a meta-analytical approach, as opposed to GOA-specific values from stock assessments etc. This also means that this is at odds with the LW parameters we use to prepare the .prm file for Atlantis, as those are, instead, taken from GOA species. I imagine this will hardly make a difference as diets are theoretical and will need to be calibrated, but it is still a corner that I am cutting. This applies to the spatial distributions as well. 
2. There is the potential to obtain very different weights by using a LW relationship compared to what was recorded as weight of the ingested prey. This has two main reasons: (a) the LW relationship is an approximation and the real weight of an individual fish may vary greatly; and (b) a prey item may have been digested. This goes two ways, in that LW may overestimate or underestimate the real weight of ingested prey. For this reason, the LW-calculated weight has the sole purpose of working out proportions of juveniles and adults in the stomach content of a predator. Those weight are not to be used further.
```{r}
length_to_weight <- function(Name,Len,lw_lmat){
  this_a <- lw_lmat %>% filter(Scientific.Name == Name) %>% pull(a)
  this_b <- lw_lmat %>% filter(Scientific.Name == Name) %>% pull(b)
  this_w <- this_a*Len^this_b
  this_w <- ifelse(length(this_w)==0, NA, this_w) # this is to keep the prey items that we do not have LW info for
  this_w # in g
}

prey_length2 <- prey_length1 %>%
  mutate(Prey_wt = purrr::pmap(list(Prey_name_new,Prey_sz_cm),possibly(length_to_weight,NA),lw_lmat_prey)) %>%
  unnest(cols = Prey_wt)
```

Group by year, **stratum**, predator name, predator stage, prey name, prey stage, and add up the weight of the ingested prey, and the relative proportions of juveniles and adults in that. We do not work at the haul level here, because of issues with the representativeness of within-haul subsamples for size compositions, and of the hauls themselves for the population size structure. Basically an issue with sample size, and with the fact that often they would find many prey items but they were impossible to size because they were fragments or were digested. Working at stratum level also aligns this better with Kerim's previous work.
```{r}
prey_length3 <- prey_length2 %>%
  group_by(Year, Stratum, Pred_name_new, Pred_nodc, Pred_stage, Prey_name_new, Prey_nodc, Prey_stage) %>%
  summarise(Prey_weight_ingested_stage = sum(Prey_wt)) %>%
  ungroup() %>%
  group_by(Year, Stratum, Pred_name_new, Pred_nodc, Pred_stage, Prey_name_new, Prey_nodc) %>%
  mutate(Prey_weight_ingested_total = sum(Prey_weight_ingested_stage),
         Prop = Prey_weight_ingested_stage/Prey_weight_ingested_total) %>% # get proportion of prey weight for each prey stage by haul
  ungroup()

# check that A+J=1 in each stratum
# prey_length3 %>% group_by(Year,Stratum,Pred_name_new,Pred_stage,Prey_name_new) %>% summarise(check = sum(Prop)) %>% na.omit() %>% pull(check) %>% mean() 
```

# Break down predator set by proportions (8)

Now we need to tie the proportion information back to the predator data set, and multiply the ingested prey **in each haul** by the proportion of each prey stage within the **stratum** where that haul is located. Where there is no stage separation for prey, the proportion will be 1, and the amount of prey is not down-scaled. Do this by haul as it is fine enough for our purposes, but bear in mind that previous work has done it by stomach. **Importantly, the size-comp information comes from the stratum level, not haul level**.

We have to join by NODC because names are inconsistent in their format between the predator and prey length sets. Be careful with handling the invertebrates here.

Missing prey stage: if the prey stage is `NA`, replace it with Undetermined (U). If the proportion for a stage is `NA`, replace it with 1. This may hide some bugs but it should correct for the (many) species that do not have life stage information based on length at maturity (e.g. invertebrates of many fish species that are not groundfish).
```{r}
predator_by_stage <- predator_by_haul %>% 
  full_join(prey_length3 %>% select(Year:Prey_stage,Prop), by = c('Year','Stratum','Pred_nodc','Pred_stage','Prey_nodc')) %>%
  rowwise() %>%
  mutate(Prey_stage = ifelse(is.na(Prey_stage),'U',Prey_stage),
         Prop = ifelse(is.na(Prop),1,Prop),
         Prey_ingested_g = Tot_weight_ingested*Prop) %>%
  ungroup()
```

# Diet by life stage (9)

Now we take the `predator_by_stage` set we just created with the ingested prey broken down by life stage, and apply the steps we developed to work out diet proportions. Before doing so, we map predators and prey species to their respective Atlantis groups by means of the NODC key. Then we need to group as appropriate (remember to maintain the stages here).

**NOTE**: here, we go back to haul level. The only step that gets done at stratum level is the stage composition of the prey. Note that using diet data at haul level may still present issues of representativeness of the sub-sample opened for prey size comp.

Break down this set into 4. Juvenile and adult predators eating juvenile and adult prey. Undetermined stages will appear in both.
```{r}
load("../data/nodc_matching_key_01012022.Rdata")

# we are missing Raja rhina (longnose skate) somehow, add it here manually).
rr <- data.frame(Name.x = 'Raja rhina', Atlantis.group = 'SKL', Nodc = 8713040108, Name.y = 'Skate_longnose', LongName = 'Longnose skate')

nodc_matching_key <- rbind(nodc_matching_key,rr)

calculate_diet_prop <- function(pred_stage,prey_stage){
 
  pred_subset <- predator_by_stage %>% 
    filter(Pred_stage %in% c(pred_stage,'U'), Prey_stage %in% c(prey_stage,'U')) %>% 
    ungroup() %>%
    select(Year:Prey_nodc,Prey_stage,Prey_ingested_g)
  
  # add the NODC matching key to map species to Atlantis groups
  pred_subset <- pred_subset %>% left_join(nodc_matching_key %>% 
                                             select(Nodc,Atlantis.group,Name.y) %>% 
                                             set_names(c("Pred_nodc","Pred_at_code","Pred_at_name")), 
                                           by = "Pred_nodc")
  # add prey now
  pred_subset <- pred_subset %>% left_join(nodc_matching_key %>% 
                                             select(Nodc,Atlantis.group,Name.y) %>% 
                                             set_names(c("Prey_nodc","Prey_at_code","Prey_at_name")), 
                                           by = "Prey_nodc")
  
  # NOTE: 15% of the prey entries are not identified (Misc teleost, misc crustacea, misc mollusca, etc.). We are losing all of those here, because they cannot be mapped to Atlantis groups (e.g., what teleost? a halibut or capelin?)
  
  pred_subset <- pred_subset %>% filter(!is.na(Prey_at_code))
  
  # Turn 'U' to J or A, to match the rest. Rationale: for predators, the results will be identical between juveniles and adults, and this is fine if no other info exists to differentiate. For prey, the value of pPREY will be the same in case of biomass pools or vertebrate prey for which we do not have size information. Bear in mind that you are double-counting all the U prey in this way, once for juvenile predators and once for adult. However, this should be alleviated by the fact that, if a life stage of the predators does not eat a certain biomass pool, this should not show when subsetting by the predator life stage. For example, assume that juvenile cod eats some copepods, and adult cod does not. Copepods are all U, but they should not show up in the adult cod data set, thus their proportion in the diet of adult cod should be 0. 
  
  pred_subset <- pred_subset %>% mutate(Pred_stage = str_replace(Pred_stage, 'U', pred_stage),
                                        Prey_stage = str_replace(Prey_stage, 'U', prey_stage))
  
  # the last step means that from here on out we can ignore the stages in the groupings, because all stages are going to be as specified in the arguments to the function call (i.e. U is no longer present)
  
  # make a counter for the number of predators aggregated by Atlantis group - which is different from the counter made by species above
  npred_ag <- pred_subset %>%
    select(Year,Hauljoin,Pred_at_name,Pred_nodc,Npred) %>%
    distinct() %>%
    group_by(Year,Hauljoin,Pred_at_name) %>%
    summarize(Npred_ag = sum(Npred)) %>%
    ungroup()
  
  # add new counter to the predator set for weighting - discard old counter used to produce this one
  pred_subset1 <- pred_subset %>% left_join(npred_ag, by = c('Year','Hauljoin','Pred_at_name')) %>%
    select(-Npred)
  
  pred_subset1 <- pred_subset1 %>%
    group_by(Year, Hauljoin, Npred_ag, Pred_at_name, Prey_at_name) %>% 
    summarize(Prey_ingested_g_1 = sum(Prey_ingested_g)) %>% # this step adds up the prey weight from species to atlantis groups
    group_by(Year, Hauljoin, Npred_ag, Pred_at_name) %>%
    mutate(Total_ingested_g = sum(Prey_ingested_g_1)) %>% # how much prey (all species) was eaten by pred Y in year Z
    group_by(Year, Hauljoin, Npred_ag, Pred_at_name, Prey_at_name) %>% 
    summarize(pcW_stom = ifelse(Total_ingested_g!=0, Prey_ingested_g_1/Total_ingested_g, 0)) %>% # proportion ingested X and total ingested
    ungroup() %>%
    complete(Prey_at_name, nesting(Year, Hauljoin, Npred_ag, Pred_at_name), fill = list(pcW_stom=0)) %>% # add zeroes, i.e. the predator had 0 g of prey XXX stage YYY in its stomach
    mutate(pcW_weighted = pcW_stom*Npred_ag) %>% # multiply the proportions by the number of predators, to give them more weight the more predators there were in a haul
    group_by(Year, Pred_at_name) %>%
    mutate(Tot_pred = sum(pcW_weighted)) %>% # this returns the total number of predators of a species in a year 
    ungroup() %>%
    group_by(Pred_at_name, Prey_at_name, Year, Tot_pred) %>% # 
    summarize(pcW_weighted = sum(pcW_weighted)/Tot_pred) %>% # sum across hauls and divide by total n of predators to have an adjusted (weighted) proportion of a prey in the diet, weighted for the number of fish in each haul
    distinct()
  
  # average over all years. Kind of along the same lines, we may want to give different weights to different years here. This may attempt to reflect the number of hauls per year, giving more weight to years that had more hauls, or perhaps we may want to dampen the effect of some atypical years. For now, do a simple mean of the proportion for each predator-prey pair. Reflect on what period you want to use for this. Using fewer years will mean loosing some species. 
  
  diet_all_years <- pred_subset1 %>% 
    group_by(Pred_at_name, Prey_at_name) %>%
    summarise(pcW_mean = mean(pcW_weighted, na.rm = TRUE)) 
  
  # ditch NaN's, which stem from a predator being recorded but not actually having a number of fished individuals - e.g. capelin (not sure why this happens in the data but there you go)
  
  diet_all_years <- diet_all_years[complete.cases(diet_all_years),]
  
  check_sum <- diet_all_years %>% group_by(Pred_at_name) %>% summarise(check = sum(pcW_mean)) %>% ungroup() %>% pull(check) %>% mean()
  
  if(check_sum!=1){warning("The diet proportions do not add up to 1")}
  
  # write this out
  write.csv(diet_all_years,paste("../output/REEM/dietprop_stratum_pred_",pred_stage,"_prey_",prey_stage,".csv",sep = ""), row.names = F)
  
  # make a plot to export
  p <- diet_all_years %>%
    mutate(pcW_mean = na_if(pcW_mean,0)) %>%
    ggplot()+
    geom_tile(aes(x = Prey_at_name, y = Pred_at_name, fill = pcW_mean), color = "darkgrey")+
    scale_fill_viridis()+
    #scale_fill_gradient2(low = "blue", high = "red", midpoint = .5, na.value="grey")+
    theme_bw()+
    theme(axis.text.x = element_text(angle = 60, hjust = 1.05, vjust = 1, size = 14),
          axis.text.y = element_text(size = 14))+
    labs(title = paste0("Proportion of predator diet (",pred_stage," predator and ",prey_stage," prey)"), 
         x = "Prey", y = "Predator",
         fill = "Proportion of ingested prey")
  
  ggsave(paste("../output/REEM/dietprop_stratum_pred_",pred_stage,"_prey_",prey_stage,".png",sep = ""),p,width = 16, height = 12)
}

# run the function and make four plots
calculate_diet_prop('J','J')
calculate_diet_prop('J','A')
calculate_diet_prop('A','J')
calculate_diet_prop('A','A')
```
